Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt

### 2024-12-09 Phase_II/draft/intro.tex
modification

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in single/binary star systems. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) range, have renewed II as a feasible technique in comparison to the limitation of amplitude interferometry for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. However, despite these technological advancements, II faces a significant limitation in terms of phase loss and the absence of complete reconstruction of the brightness distribution of stellar sources.

@@

Later \cite{holmes2010two} proposed an alternative method, utilizing the Cauchy-Riemann relations to reconstruct 1-D images and extending this to 2-D images across a range of signal-to-noise (SNR) values. This algorithm was applied to simulated data of stellar objects using II, considering existing and forthcoming Imaging Cherenkov Telescope Arrays (ICTAs) with a large number of telescopes \citep{nunez2010stellar, nunez2012high, nunez2012imaging}. Despite these theoretical advances, practical application to actual observational data of stellar objects with II has yet to be achieved. There is a need to advance this field by introducing a novel image reconstruction technique.

Currently, artificial intelligence (AI) is applied to high-volume data to reduce human computing resources and time effort. New analysis techniques using Machine Learning (ML) applications on observed data are playing at the front to reveal the astronomical mystery. In radio morphologies, the supervised and unsupervised algorithm is used to distinguish the radio components from the infrared source \citep{wu2019radio, galvin2020cataloguing}. Recently, the image reconstruction of the M87 black hole using Generative Adversarial Networks (GANs) has been performed \citep{10.1093/mnras/stad3797}. There is also a need to introduce the application of ML to II data, which will aim to overcome the limitations of image reconstruction and provide high-resolution optical astronomy.

In this paper, we propose an alternative approach for image reconstruction of stellar objects with II using GAN \citep{goodfellow2014generative}. We consider four ICTAs and perform a one-night simulation for a fast-rotating star. The predicted image by a trained GAN shows promising results for reconstructing the image's shape and size. Image moments evaluate the predicted image and demonstrate that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only up to the third-order moment. We expect even greater precision of image reconstruction by including more telescopes, especially to recover the higher-order image moments. It holds great promise, pushing the boundaries of high-resolution optical astronomy and opening up new possibilities for studying stellar objects with remarkable detail.

This paper is organized as follows: The first section discusses Intensity Interferometry following its signal and noise for fast-rotating stars along the Earth's rotation. The second section introduces the GAN formulation and its structure. The third section details the parameter selection for training the GAN to reconstruct the image. After that, the fourth section presents the results of trained GAN visually and using image moments. At the end, there is a conclusion of the entire result.

>>>>

Intensity Interferometry (II) emerged in the 1970s as a groundbreaking technique for measuring the diameters, orbits, and limb-darkening coefficients of 32 stars in single/binary/multiple star systems \cite{brown1974intensity}. Despite these significant achievements, the method did not gain widespread adoption due to the then-required highly sensitive photon detectors and sophisticated data analysis techniques. However, advancements in computational methods and modern electronics, which now offer time resolution of detectors in the nanoseconds (ns) range, have renewed II as a feasible technique in comparison to the limitation of amplitude interferometry for resolving astrophysical objects at visible wavelengths. Current simulations with ns resolution of photon detectors show that II is effective in achieving high precision measurement of parameters for stellar objects \citep{10.1093/mnras/stab2391, 10.1093/mnras/stac2433}. However, despite these technological advancements, II faces a significant limitation in terms of phase loss and the absence of complete reconstruction of the brightness distribution of stellar sources.

@@

Later \cite{holmes2010two} proposed an alternative method, utilizing the Cauchy-Riemann relations to reconstruct 1-D images and extending this to 2-D images across a range of signal-to-noise (SNR) values. This algorithm was applied to simulated data of stellar objects using II, considering existing and forthcoming Imaging Cherenkov Telescope Arrays (ICTAs) with a large number of telescopes \citep{nunez2010stellar, nunez2012high, nunez2012imaging}. Despite these theoretical advances, practical application to actual observational data of stellar objects with II has yet to be achieved. The introduction of a novel image reconstruction technique with II will advance the field of optical astronomy.

Currently, Artificial Intelligence (AI) is applied to high-volume data to reduce human computing resources and time effort. New analysis techniques using Machine Learning (ML) applications on observed data are playing at the front to reveal the astronomical mystery. In radio morphologies, the supervised and unsupervised algorithm is used to distinguish the radio components from the infrared source \citep{wu2019radio, galvin2020cataloguing}. Recently, the image reconstruction of the M87 black hole using Generative Adversarial Networks (GANs) has been performed \citep{10.1093/mnras/stad3797}. There is also a need to introduce the application of ML to II data, which will aim to overcome the limitations of image reconstruction and provide high-resolution optical astronomy.

In this paper, we propose an alternative approach for image reconstruction of stellar objects with II using conditional GAN (cGAN) \citep{goodfellow2014generative}. We consider four ICTAs and perform a one-night simulation for a fast-rotating star. The predicted image by a trained GAN shows promising results for reconstructing the image's shape and size. Image moments evaluate the predicted image and demonstrate that the size and shape of the fast-rotating star are reconstructed. However, the brightness distribution recovers only up to the third-order moment. We expect even greater precision in image reconstruction by including more telescopes, especially in recovering higher-order image moments. It holds great promise, pushing the boundaries of high-resolution optical astronomy and opening up new possibilities for studying stellar objects with remarkable detail.

This paper is organized as follows: The first section discusses Intensity Interferometry following its signal and noise for fast-rotating stars along the Earth's rotation. The second section introduces the GAN formulation and its structure. The third section details the parameter selection for training the GAN to reconstruct the image. After that, the fourth section presents the results of trained GAN visually using image moments. At the end, there is a conclusion of the entire result.

### 2024-12-09 Phase_II/draft/main.tex
modification

\begin{abstract}
In modern astronomy, simulation and observation with Intensity Interferometry (II) are showing noteworthy progress with optical wavelengths to achieve high-resolved stellar objects. Despite these achievements, phase retrieval of the signal remains challenging due to photon correlation, hence the image reconstruction. Here, we propose the application of a conditional Generative Adversarial Network (cGAN) to II simulated data to overcome the limitations of traditional image reconstruction methods, as Artificial Intelligence (AI) techniques are at the forefront of revealing astronomical mysteries by analyzing the signal. Here, the cGAN effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by utilizing sparsely sampled, phase-subtracted Fourier transforms of the star. Simulations of II are based on an assembly of four telescopes similar to existing arrays. However, the sensitivity of the signal and high resolution are expected to improve with additional baselines, which makes the current and future Cherenkov Telescope Array Observatory (CTAO) an ideal candidate for II applications. This approach is highly relevant to the existing image reconstruction technique with II and could revolutionize high-resolution imaging in astronomy.
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

>>>>

\begin{abstract}
In modern astronomy, simulation and observation with Intensity Interferometry (II) are showing noteworthy progress with optical wavelengths to achieve high-resolved stellar objects. Despite these achievements, phase retrieval of the signal remains challenging due to photon correlation, hence the image reconstruction. Here, we propose the application of a conditional Generative Adversarial Network (cGAN) to II simulated data to overcome the limitations of traditional image reconstruction methods, as Artificial Intelligence (AI) techniques are at the forefront of revealing astronomical mysteries by analyzing the signal. Here, the cGAN effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by utilizing sparsely sampled, phase-subtracted Fourier transforms of the star. This approach is highly relevant to the existing image reconstruction technique with II and could revolutionize high-resolution imaging in astronomy. Simulations of II are based on an assembly of four telescopes similar to existing arrays. However, the sensitivity of the signal and resolution of stellar objects are expected to improve with additional baselines, which makes the current and future Cherenkov Telescope Array Observatory (CTAO) an ideal candidate for II applications.
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}
